\name: A Non-Parametric, Hierarchical \\ Wavelength-Calibration Model for a Precision Spectrograph

### 2020-05-29 paper/paper.tex
Added Needed Number of Exposures

\begin{figure*}[h]
\centering
\includegraphics[width=\textwidth]{Figures/methodCartoon.pdf}
\caption{A cartoon representation of the \name\ method, as described in Section \ref{sec:method}.  We enlarge variations in measured line position, changes in calibration space, and interpolation deviations for clarity.  In step one, dimensionality reduction and denoising (\textsection \ref{sec:denoising}), the complete set of line positions for all exposures is analyzed to return a set of $K$ basis vectors, $g_n$, which represent different ways the spectrograph calibration changes.  These basis vectors span the $K$-dimensional calibration space of the spectrograph, which includes all possible wavelength solutions.  In step two (\textsection \ref{sec:interp_time}), \name\ interpolates the amplitudes of each basis vector, $a_n,k$, to return the calibration state for a specific science exposure, returned as a set of de-noised calibration lines.  The assigned wavelengths of these de-noised line positions are then interpolated onto other pixels in step three (\textsection \ref{sec:interp_wsol}) to construct a full wavelength model that returns wavelength as a function of detector position $x$ and echelle order $m$.}
\label{fig:cartoon}
\end{figure*}

@@

@@

\begin{figure}[h]
\centering
\includegraphics[width=.6\textwidth]{Figures/all_results.pdf}
\caption{Difference in predicted and theoretical wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.  The per line RMS as defined in Equation \ref{eq:rms} is given in the top-right corner in each method's corresponding color.  \name\ returns the smallest spread in residuals.}
\label{fig:testHists}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=.5\textwidth]{Figures/kvals_all.pdf}
\caption{Per-line RMS of returned wavelength models for different values of $K$.  Per-line RMS is defined in Equation \ref{eq:rms} and provides a measure of the accuracy of a wavelength model.  There is a dotted, vertical line at $K=6$, our chosen $K$ value.  The dotted, horizontal line is the RMS for $K=6$.  The improvement at around $K=6$ plateaus.}
\label{fig:kvals}
\end{figure}

@@

Figure \ref{fig:pcLfc} shows the fiducial calibration state and the first 9 principal components constructed using LFC lines, which represent deviations from the fiducial calibration state.  There is clear structure in the first and second principal components.  Components three through six show smaller or more localized structure.  Components three and four have aberrant behavior on the edges of the two bluest echelle orders, where lower signal results in more variation in the line fits.  Later principal components become dominated by noise and show less coherent structure.

\subsection{Interpolation of Calibration State with Respect to Time}
\label{sec:choice_avt}
Figure \ref{fig:nightlyVariation} shows the amplitude of the first and second principal component with respect to time on the left.  Though there exists a complex overall shape to the amplitudes with respect to time, a clear linear trend exists within each night.  This is shown by the right plots in Figure  \ref{fig:nightlyVariation}.  As the beginning-of-night and end-of-night calibration sets always include LFC exposures, we use a simple linear interpolation to interpolate principal component amplitudes with respect to time.

\begin{figure*}[b]
\centering
\includegraphics[width=\textwidth]{Figures/pcAs_byDay.png}
\caption{Amplitude of the first two principal components shown as a function of time (left) or hour from midnight (right).  The top row of plots shows the amplitudes for the first principal component while the bottom row shows the amplitudes for the second principal component.  Lines show the result of a linear interpolation.  In the top, right plot, the temperature of the optical bench is also plotted in orange.  In the right plot, the principal component amplitudes have been artificially offset by the median amplitude for each night.  All days are therefore roughly on the same scale, but the y-axis is different from the left plots.  In the right column, points are colored by the MJD of each exposure.}
\label{fig:nightlyVariation}
\end{figure*}

@@

\begin{figure*}[h]
\centering
\includegraphics[width=\textwidth]{Figures/intpx_tests.png}
\caption{Residuals from different interpolation schemes over pixels in echelle order 100.  ThAr lines, shown as blue circles, are used to construct a wavelength solution that is then evaluated at each LFC line, shown as blue lines.  The residuals of each wavelength solution for a subset of the order is shown on the left.  Histograms of the residuals for each method for the complete order is shown on the right.  Here, the y-axis is zoomed in to show the shape of each histogram more clearly.  Note: there is a blended ThAr line at approximately pixel 5300, the last ThAr line plotted.}
\label{fig:xinterp}
\end{figure*}

@@

To determine the needed line density, we systematically removed LFC lines from the \expres\ data and calculated the RMS of the returned wavelengths for the removed lines.  Figure \ref{fig:density} plots the scatter of residuals of the wavelengths returned by \name\ as a function of separation between lines in units of resolution element.  It gives an approximate estimation of the expected error between lines of different separation.  Note, however, that the stability of the lines and their measured line centers quickly becomes the dominant error source in calibration lamps over the separation between lines.  The needed line density depends on the resolution of a given spectrograph and the needed precision of the returned wavelength solution.

@@

\lz{I completely made up the below paragraph...}

The number of calibration exposures needed is linked to the amount of variability in the instrument.  Calibration exposures must span the range of calibration space accessible by the spectrograph, or at least accessed by the science calibrations that require wavelength solutions.  Otherwise, the dimensionality-reduction method will not have the information required to construct a basis vector that spans the entire accessible calibration space.  More observations are always better to help de-noise the calibration data.

>>>>

\begin{figure*}[t]
\centering
\includegraphics[width=\textwidth]{Figures/methodCartoon.pdf}
\caption{A cartoon representation of the \name\ method, as described in Section \ref{sec:method}.  We enlarge variations in measured line position, changes in calibration space, and interpolation deviations for clarity.  In step one, dimensionality reduction and denoising (\textsection \ref{sec:denoising}), the complete set of line positions for all exposures is analyzed to return a set of $K$ basis vectors, $g_n$, which represent different ways the spectrograph calibration changes.  These basis vectors span the $K$-dimensional calibration space of the spectrograph, which includes all possible wavelength solutions.  In step two (\textsection \ref{sec:interp_time}), \name\ interpolates the amplitudes of each basis vector, $a_n,k$, to return the calibration state for a specific science exposure, returned as a set of de-noised calibration lines.  The assigned wavelengths of these de-noised line positions are then interpolated onto other pixels in step three (\textsection \ref{sec:interp_wsol}) to construct a full wavelength model that returns wavelength as a function of detector position $x$ and echelle order $m$.}
\label{fig:cartoon}
\end{figure*}

@@

\begin{figure}[t]
\centering
\includegraphics[width=.6\textwidth]{Figures/all_results.pdf}
\caption{Difference in predicted and theoretical wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.  The per line RMS as defined in Equation \ref{eq:rms} is given in the top-right corner in each method's corresponding color.  \name\ returns the smallest spread in residuals.}
\label{fig:testHists}
\end{figure}

@@

@@

\begin{figure}[h]
\centering
\includegraphics[width=.5\textwidth]{Figures/kvals_all.pdf}
\caption{Per-line RMS of returned wavelength models for different values of $K$.  Per-line RMS is defined in Equation \ref{eq:rms} and provides a measure of the accuracy of a wavelength model.  There is a dotted, vertical line at $K=6$, our chosen $K$ value.  The dotted, horizontal line is the RMS for $K=6$.  The improvement at around $K=6$ plateaus.}
\label{fig:kvals}
\end{figure}

@@

Figure \ref{fig:pcLfc} shows the fiducial calibration state and the first 9 principal components constructed using LFC lines, which represent deviations from the fiducial calibration state.  There is clear structure in the first and second principal components.  Components three through six show smaller or more localized structure.  Components three and four have aberrant behavior on the edges of the two bluest echelle orders, where lower signal results in more variation in the line fits.  Later principal components become dominated by noise and show less coherent structure.
\pagebreak
\subsection{Interpolation of Calibration State to Science Exposures}
\label{sec:choice_avt}
Figure \ref{fig:nightlyVariation} shows the amplitude of the first and second principal component with respect to time on the left.  Though there exists a complex overall shape to the amplitudes with respect to time, a clear linear trend exists within each night.  This is shown by the right plots in Figure  \ref{fig:nightlyVariation}.  As the beginning-of-night and end-of-night calibration sets always include LFC exposures, we use a simple linear interpolation to interpolate principal component amplitudes with respect to time.

\begin{figure*}[t]
\centering
\includegraphics[width=\textwidth]{Figures/pcAs_byDay.png}
\caption{Amplitude of the first two principal components shown as a function of time (left) or hour from midnight (right).  The top row of plots shows the amplitudes for the first principal component while the bottom row shows the amplitudes for the second principal component.  Lines show the result of a linear interpolation.  In the top, right plot, the temperature of the optical bench is also plotted in orange.  In the right plot, the principal component amplitudes have been artificially offset by the median amplitude for each night.  All days are therefore roughly on the same scale, but the y-axis is different from the left plots.  In the right column, points are colored by the MJD of each exposure.}
\label{fig:nightlyVariation}
\end{figure*}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=\textwidth]{Figures/intpx_tests.png}
\caption{Residuals from different interpolation schemes over pixels in echelle order 100.  ThAr lines, shown as blue circles, are used to construct a wavelength solution that is then evaluated at each LFC line, shown as blue lines.  The residuals of each wavelength solution for a subset of the order is shown on the left.  Histograms of the residuals for each method for the complete order is shown on the right.  Here, the y-axis is zoomed in to show the shape of each histogram more clearly.  Note: there is a blended ThAr line at approximately pixel 5300, the last ThAr line plotted.}
\label{fig:xinterp}
\end{figure*}

@@

Implementing \name\ will require an existing library of calibration exposures that span the range of calibration space accessible by the spectrograph, or at least the space accessed by the science exposures being calibrated.  A new set of calibration exposures would be needed for any changes to the instrument that could significantly change the calibration space.  While there is no exact cutoff for how many exposures are needed, the number is certainly much larger than $K$, the dimensionality of the calibration space.  More calibration exposures will help with denoising .  The number of calibration exposures required throughout a night will depend on how much the instrument varies throughout the night, and the chosen interpolation scheme, as mentioned in Section \ref{sec:choice_avt}.

As an empirical assessment of the needed line density, we systematically removed LFC lines from the \expres\ data and calculated the per-line RMS of the returned wavelengths for the removed lines.  Figure \ref{fig:density} plots the scatter of residuals of the wavelengths returned by \name\ as a function of separation between lines in units of resolution element.  It gives an approximate estimation of the expected error between lines of different separation.  Note, however, that the stability of the lines and their measured line centers quickly becomes the dominant error source in calibration lamps over the separation between lines.  The needed line density depends on the resolution of a given spectrograph and the needed precision of the returned wavelength solution.

@@
